Network Deliberation: The role of network structure in large-scale, internet-enabled, participatory decision-making

### 2022-05-05 UMich/chapters/ch01-intro.tex
Import NetDelibExp and formatting fixes.
Update intro.

In recent years, examples of large-scale collaborations have emerged that seem
to achieve the previously unachievable. Millions of volunteer Wikipedia editors
have created a high-quality encyclopedia without centralized leadership
\cite{keegan_evolution_2017, giles_internet_2005}.
The Free Software movement has produced the Linux kernel and GNU operating
system, which power much of the modern internet
\cite{coleman_coding_2013, benkler_coases_2002, raymond_cathedral_1999}.
Social movements such as the Arab Spring, Occupy, Black Lives Matter, and
Podemos have reshaped national governments and brought attention to deeply
entrenched social issues
\cite{tufekci_twitter_2017, gonzalez-bailon_networked_2016}.
The emergence of these large-scale decentralized collaborations has been
attributed to the fast, bidirectional,
and global communication enabled by the internet
\cite{tufekci_twitter_2017, benkler_coases_2002}.
A better understanding of how specifically such communication sidesteps
historical barriers to large-scale collaboration will contribute to more
effective policy as well as best practices for organizational design and
intervention.
This dissertation focuses on one particularly challenging aspect of
such collaborations: decision-making
I specificaly examine how,
when groups are too large for all members to participate in all discussions,
the course and outcome of the decision-making process is influenced by the
communication network structure: the shape of who talks to whom.

@@

To bridge the limitations of the above methods,
this dissertation will include a controlled field experiment evaluating the
effect of network deliberation in real-world collective decision-making
(Chapter \ref{chap:experiment}).

@@

Chapter \ref{chap:experiment} proposes and describes current progress on a
field experiment to study network deliberation in large-scale human collaborations.
The study design uses periodic ranked-choice polls to track individual preferences
throughout the course of a large-scale online deliberation.
By varying communication network structure and tracking the evolution of individual
preferences,
this experiment evaluate the ability of network deliberation to resolve conflict
and build consensus, relative to conventional single-group deliberation.

\section{Dissertation Progress and Timeline}
At the time of this proposal,
the analysis of English-language WikiProjects described in Chapter
\ref{chap:wp-prod-perf} has been completed and resulted in a published
paper \cite{platt_network_2018}.
The agent-based modelling project described in Chapter \ref{chap:abm}
has been completed and resulted in a finished manuscript.
I have completed custom software, obtained IRB exemption, and conducted
several pilot studies for the field experiment described in
Chapter \ref{chap:experiment}.
I expect to have data collected from at least one full-scale experiment
by the end of 2021, and to have the data analyzed by mid-February, 2022.
I plan to have my dissertation completed by mid-May, 2022.

>>>>

In recent years, examples of large-scale collaborations have emerged that seem
to achieve the previously unachievable. Millions of volunteer Wikipedia editors
have created a high-quality encyclopedia without centralized leadership
\cite{keegan_evolution_2017, giles_internet_2005}.
The Free Software movement has produced the Linux kernel and GNU operating
system, which power much of the modern internet
\cite{coleman_coding_2012, benkler_coases_2002, raymond_cathedral_1999}.
Social movements such as the Arab Spring, Occupy, Black Lives Matter, and
Podemos have reshaped national governments and brought attention to deeply
entrenched social issues
\cite{tufekci_twitter_2017, gonzalez-bailon_networked_2016}.
The emergence of these large-scale decentralized collaborations has been
attributed to the fast, bidirectional,
and global communication enabled by the internet
\cite{tufekci_twitter_2017, benkler_coases_2002}.
A better understanding of how specifically such communication sidesteps
historical barriers to large-scale collaboration will contribute to more
effective policy as well as best practices for organizational design and
intervention.
This dissertation focuses on one particularly challenging aspect of
such collaborations: decision-making
I specificaly examine how,
when groups are too large for all members to participate in all discussions,
the course and outcome of the decision-making process is influenced by the
communication network structure: the shape of who talks to whom.

@@

To bridge the limitations of the above methods,
this dissertation includes a controlled field experiment evaluating the
effect of network deliberation in real-world collective decision-making
(Chapter \ref{chap:experiment}).

@@

Chapter \ref{chap:experiment} describes an experimental study network deliberation in large-scale human collaborations.
The study design uses periodic ranked-choice polls to track individual preferences
throughout the course of a large-scale online deliberation.
By varying communication network structure and tracking the evolution of individual
preferences,
this experiment evaluates the ability of network deliberation to resolve conflict
and build consensus, relative to conventional single-group deliberation.
The primary contributions are:
\begin{itemize}
\setlength\itemsep{0pt}
    \item Findings support the hypothesis that network deliberation is better at facilitating agreement than conventional deliberation.
    \item Findings suggest that network deliberation can provide protection against information cascades.
    \item Findings include no evidence that network deliberation facilitates substantial conflict-resolution, but indicate that it may provide protection against polarization.
\end{itemize}
